---
title: "Basic Usage"
output: 
  rmarkdown::html_vignette:
      toc: true
vignette: >
  %\VignetteIndexEntry{fiphde}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval=FALSE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r, eval=TRUE}
knitr::knit_exit()
```


## Overview

_Fixme intro text here..._

## Usage

First, load the `fiphde` package and other ancillary packages.

```{r setup}
library(fiphde)
library(fiphde)
library(dplyr)
library(purrr)
library(readr)
library(ggplot2)
library(furrr)
```

### Get data

First, let's retrieve Influenza-like Illness (ILI) data from [CDC FluView](https://gis.cdc.gov/grasp/fluview/fluportaldashboard.html). Note that `years` argument gets the *flu season* corresponding to the year. If you only get 2020-2021, you're getting the 2020-2021 and 2021-2022 flu season, meaning that you will *NOT* capture early 2020 data.

```{r, message = FALSE}
ilidat <- get_cdc_ili(region=c("national","state"), years=2019:lubridate::year(lubridate::today()))
```

Some states have missing ILI data so we pull from [CMU Delphi NowCast](https://delphi.cmu.edu/nowcast/) instead.

```{r, message = FALSE}
# Replace FluView data with NowCast data for Florida only
ilidat <- state_replace_ili_nowcast_all(ilidat, state="FL")

## Pull NowCast data for all states for a specified date
iliaug <- replace_ili_nowcast(ilidat, weeks_to_replace=1)
```

### Fit ARIMA model

Now that the data has been gathered we being model fitting. We use `forecast_ili` to fit a `fable::ARIMA` model with parameters optimized automatically. Here 

```{r, message = FALSE}
ilidat_st <- iliaug %>% dplyr::filter(region_type=="States")
ilifor_st <- forecast_ili(ilidat_st, horizon=4L, trim_date="2020-03-01")
```
